% !TeX root = ./report.tex
\chapter{Approach}
The approach was set out by first analysing existing methods of preserving annotations through intermediate representations. These include the:
\begin{itemize}
    \item the compCert Verified C Compiler, 
    \item GNU Extension for Extended Inline Assembly, and
    \item modifying the LLVM compiler to preserve annotations throughout intermediate representations.
\end{itemize}

Each of these approaches will be analysed individually for viability across each of the test cases outlined in section \ref{sec:testCases}. For approaches that pass all necessary test cases, a further analysis will be conducted into its suitability and development of any necessary tools to assist in the preservation technique, as outlined in sections \ref{sec:qualityAnalysis} and \ref{sec:toolDevelopment}. Finally, an analysis on the runtime efficiency of the program will be conducted to assess the success of the annotations with various levels of optimisation. The approach for this analysis is outlined in \ref{sec:runtimeEfficiencyOptimisation}.

\section{Test Cases}
\label{sec:testCases}
A suite of test C programs (See Appendix \ref{app:testPrograms}) were created to assist in guiding the process of evaluating each approach as a possible means of preserving annotations. Each program has inline comments documenting the annotation that should be preserved and its location within the program. Additionally, each program aims to test a separate element required to perform a static wpif analysis. Namely, these are to preserve the following through to the assembly output:

\begin{enumerate}
    \item comments,
    \item simple and complex variables (e.g. struct elements and volatile global variables),
    \item security policies,
    \item predicates on the initial state, and
    \item loop invariants.
\end{enumerate}

Each test was conducted to assess the viability of each approach of preserving annotations. If the approach cannot preserve all the required annotations described in the aforementioned list, then it is not viable for a wpif analysis and another technique must be explored.

The justification for each of the test files are as follows:

\miniheading{comment.c}
This test case is primarily a stepping stone to testing more complex scenarios. Here we have a generic comment "critical comment" and we are looking to preserve it through to the assembly. As well as preserving the comment itself, the location of the comment within the source code is to be preserved.

\miniheading{variable.c}
The test file \textit{variable.c} builds off \textit{comment.c}, however, we are additionally looking to preserve annotations about local variables within the program. Here multiple variable types are tested:

\begin{itemize}
    \item int,
    \item char,
    \item unsigned int,
    \item short,
    \item long,
    \item float, and
    \item double.
\end{itemize}

With each of these variables their type data is included as an annotation. This test is particularly interesting as with higher levels of optimisation we can observe how the annotations behave when a variable is optimised out.

\miniheading{volatile.c}
This test program looks at how the technique handles volatile variables. A variable declared as volatile tells the program its value could change unexpectedly. This is especially important when dealing with concurrent programs. If the technique cannot handle volatile variables it is unable to be used for a wpif analysis.

\miniheading{loop.c}
This test program tests how the annotator handles loops and loop invariants. It contains security policies, predicates on the initial state and loop invariants.

\miniheading{rooster.c}
The test program, \textit{rooster.c} delves into a more complex program, combining several features of the previous tests. It contains annotations within functions and global variables.

\miniheading{password.c}
This program tests how annotations are preserved within structs, a user-defined data type. Additionally, \textit{password.c} is a more complex program with multiple functions.

\miniheading{deadStoreElimination.c}
Testing dead store elimination is a bit more complex, as it requires comparing the compiled output before and after compiler optimisations are turned on. Here, the test program simulates the program described in section \ref{sec:compilersSecurity}.

\miniheading{pread.c}
The program \textit{pread.c} is a culmination of all the previous test cases, and is similar to \textit{loop.c}, however, the global variables within it are volatile. It requires all the necessary components for a wpif analysis.

\section{Quality Analysis}
\label{sec:qualityAnalysis}
Although a method of preserving c annotations may be able to successfully pass all the test cases, it is important to avoid modifying the assembly instructions. The reason for performing a static analysis on the compiled output is due to the optimisations performed by the compiler. As such, it is important to ensure that preserved annotations do not remove or undo any optimisations that may have been performed by the compiler. 

The methodology for testing quality in this manner is to compare the compiled assembly output for a program with annotations on to the compiled assembly output for the same program without annotations. If unnecessary assembly instructions have been added, it is indicative that the annotations has modified the program in unintended ways. 

\section{Efficiency and Optimisation}
\label{sec:runtimeEfficiencyOptimisation}
In the case that the annotations have introduced additional statements into the compiled assembly output, understanding the extent of these changes is important. Here, a efficiency analysis can be conducted on the assembly. Using big O notation, an upper bound can be placed on the program. Doing so allows for a comparison of the efficiency of the annotated and non-annotated assembly. 

Let \(A(n)\) be the function describing the annotated assembly, and \(B(n)\) be the function describing the non-annotated assembly. Then,

\[A(n) \in \Theta(g(n))\]
\[B(n) \in \Theta(h(n))\]

If the non-annotated assembly has a lower bound than the annotated assembly, such that

\[h(n) \in O(g(n)), and\]
\[g(n) \notin O(h(n))\]

then the annotations have modified the program in a way that reduces runtime efficiency. It is important to detect when this has happened, as it indicates the annotations have reversed the intended compiler optimisations.

In the case where the annotation process has resulted in additional assembly instructions inserted into the compiled output, however, they do not reduce runtime efficiency in terms of big O notation, a empirical analysis of the runtime duration of a program can be conducted to assess the disadvantage of the annotated program.

\section{Tool Development}
\label{sec:toolDevelopment}
In cases where it is appropriate, a tool may be developed to assist in the annotating process. This tool may either:
\begin{itemize}
    \item assist in the annotating process,
    \item verify the correctness of the annotations, or
    \item perform additional analysis on the compiled output.
\end{itemize}
If the approach of modifying the LLVM compiler is pursued, developing such a tool to assist the annotating process will be necessary.