\documentclass[12pt, a4paper]{report}
\usepackage[utf8]{inputenc}
\usepackage{biblatex}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{xcolor}
\usepackage[toc,page]{appendix}
\usepackage{setspace}
\usepackage[margin=1in]{geometry}
\setstretch{1.5}
\setlength{\parindent}{0em}
\setlength{\parskip}{1em}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=true,
    breaklines=true, 
    captionpos=b,
    keepspaces=true, 
    numbers=left,
    numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=2
}
\lstset{style=mystyle}

\lstdefinelanguage
   [x64]{Assembler}     % add a "x64" dialect of Assembler
   [x86masm]{Assembler} % based on the "x86masm" dialect
   % with these extra keywords:
   {morekeywords={CDQE,CQO,CMPSQ,CMPXCHG16B,JRCXZ,LODSQ,MOVSXD, %
                  POPFQ,PUSHFQ,SCASQ,STOSQ,IRETQ,RDTSCP,SWAPGS, %
                  rax,rdx,rcx,rbx,rsi,rdi,rsp,rbp, %
                  r8,r8d,r8w,r8b,r9,r9d,r9w,r9b, %
                  r10,r10d,r10w,r10b,r11,r11d,r11w,r11b, %
                  r12,r12d,r12w,r12b,r13,r13d,r13w,r13b, %
                  r14,r14d,r14w,r14b,r15,r15d,r15w,r15b}} % etc.

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\miniheading}[1]{\noindent\makebox[\columnwidth][l]{\textbf{#1}}}

\setlength{\columnsep}{15pt}

\graphicspath{ {./images/} }
\addbibresource{bibliography.bib}

\title{C Annotations for Concurrent Information Flow Security}
\author{Alexander Blyth}
\date{June 2021}

\begin{document}
\maketitle

\tableofcontents

\clearpage
\begin{center}
    \large{\textbf{Abstract}}
\end{center}
The aim of this thesis is to to preserve annotations written in C source code through intermediate representations of compilation and compiler optimisations to assist verification of information flow security in concurrent programs through static analysis techniques. Information flow security in concurrency is difficult due to the increasing complexity introduced with multiple threads. Additionally, compiler optimisations can break security guarantees that have been verified in source code. Thus, any verification of security policies must be performed at assembly or lower representations. Three methods of preserving annotations were analysed for assessing the suitability  for verifying information flow security in concurrent programs. Each technique was assessed to observe if all necessary annotations can be preserved and to observe how it performs through aggressive compiler optimisations. CompCert AIS annotations were robust and preserved annotations through aggressive compiler optimisations, however, it does not support volatile variables which are essential for concurrent programs. CompCert builtin annotations do allow annotations of volatile variables, however, it is not robust and compiler optimisations were not always performed due to the annotations. Finally, inline assembly was excellent at preserving annotations, and a tool was developed to assist annotating assembly using this method.

\chapter{Topic Definition}
There is a high degree of complexity in verifying security guarantees in concurrent programs \cite{mantel2014noninterference}\cite{smith2019value}\cite{vaughan2012secure}. Additionally, aggressive compiler optimisations can modify the binary output in unexpected ways \cite{d2015correctness}. To preserve the security of a program, the flow of sensitive information must be protected to avoid flowing in to untrusted sources \cite{balliu2014logics}. This is where static analysis tools can be used to verify the integrity of security guarantees and the flow of sensitive information. In this thesis, we look to explore a solution to information flow security in concurrent programs through analysing the output after aggressive compiler optimisations.

We propose a tool to analyse C programs to detect security violations in information flow control. This tool will preserve annotations provided by the programmer in source code through lowering passes and aggressive compiler optimisations. The tool will work alongside the \textit{Weakest Precondition for Information Flow} (wpif) transformer described by Winter et al. \cite{winter2020information} to allow the programmer to assess the security of information flow in their concurrent programs.

Similar tools for propagating annotations and properties through compiler optimisations have been explored \cite{vu2020secure} \cite{schommer2018embedded} \cite{leroy2016compcert}, however, these tools focus on either generic solutions for propagating properties or to assist the static analysis of the \textit{Worst Case Execution Time}.

\input{background.tex}

\clearpage
\input{approach.tex}
\input{execution.tex}
\input{conclusion.tex}

% Recommendations
% - Delve into modifying compiler techniques
% - 
\clearpage
\printbibliography
\clearpage

\input{appendix.tex}
\end{document}
